---
title: "ST558 Homework 3"
author: "Kyra Kapsaskis"
format: html
editor: visual
---

## Task 1

1.  If your working directory is `myfolder/homework/`, what relative path would you specify to get the file located at `myfolder/MyData.csv`?

> `my_data <- read_csv("MyData.csv")`

2.  What are the benefits of using R projects?

> It uses the same working directory and it keeps a history of your stored objects and files in your directory folder.

3.  What is git and what is github?

> Git allows you to create repositories and track the changes to a project. Github is the portal that allows people to see your work and interact with it.

4.  What are the two main differences between a `tibble` and a `data.frame`?

> Tibble is much neater and more concise when printing. It includes some useful information that a data frame does not. And not all operations can be done on all of the columns in a tibble. There are more restrictions for doing math on the tibble.

5.  Rewrite the following nested function call using baseR's chaining operator:

```{r}
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length <
1.55), Species)
```

```{r}
iris |>
  arrange(Species) |>
  filter(Petal.Length < 1.55) |>
  select(starts_with("Petal"), Species) 

```

6.  What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

> Longer usually has unwanted variables or unnecessary NA values removed. Wider has the potential to create missing values. Both have their benefits for different statistical goals and types of data.

## Task 2

```{r}
install.packages("tidyverse")
```

```{r}
library(tidyverse)
```

### Glass data

1.  Read the data into R (from URL, using tidyverse functions, add column names)

```{r}
# I decided to use the col_names argument to create the column names

glass_data <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/glass.data", col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe", "Type of Glass"))
```

-   print out the tibble by calling the object name

```{r}
# viewing the column names to make sure they worked correctly
print(glass_data)
```

2.  use the mutate() function to start a chain that overwrites the Type_of_glass variable.
    -   use the ifelse() function to create a character string version

```{r}
glass_data <- glass_data |> 
  
  mutate(`Type of Glass` = ifelse(`Type of Glass` == 1, "building_windows_float_processed", ifelse(`Type of Glass` == 2, "building_windows_non_float_processed", ifelse(`Type of Glass` == 3, "vehicle_windows_float_processed", ifelse(`Type of Glass` == 4, "vehicle_windows_non_float_processed", ifelse(`Type of Glass` == 5, "containers", ifelse(`Type of Glass` == 6, "tableware", ifelse(`Type of Glass` == 7, "headlamps", "NA")))))))
  ) |>
  filter(Fe < 0.2 & (`Type of Glass` == "tableware" | `Type of Glass` == "headlamps"))

```

```{r}
# checking the new names and filters
print(glass_data)
```

### Yeast Data

1.  Read the data into R (same process as above, make sure to check delimiter)
    -   add column names
    -   print out the tibble

```{r}

# need to use read_table() because the delimiter is a space

yeast_data <- read_table("https://www4.stat.ncsu.edu/~online/datasets/yeast.data", col_names = c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox", "vac", "nuc", "class"))
```

2.  start a chain that removes the seq_name and nuc columns

```{r}
yeast_data <- yeast_data |>
  select(mcg:vac, class)
```

```{r}
print(yeast_data)
```

3.  continue chain. add columns corresponding to the mean and median of each numeric variable at each class grouping (see the across() function as we did in the dplyr video)

```{r}
yeast_data <- yeast_data |>
  select(mcg:vac, class) |>
  group_by(class) |>
  mutate(across(mcg:vac, mean, .names = "mean_{.col}")) |>
  mutate(across(mcg:vac, median, .names = "median_{.col}"))
```

## Task 2 - combining excel and delimited data

### reading in the white wine data

```{r}
library(readxl)
white_wine <- read_excel("white-wine.xlsx")
print(white_wine)
```

```{r}
# creating a new object with the correct variable names
white_wine_variables <- read_excel("white-wine.xlsx", 
    sheet = "variables")
print(white_wine_variables)
```

```{r}
#converting the column names
colnames(white_wine) <- c(white_wine_variables$Variables)
```

```{r}
# How does it look?

print(white_wine)
```

```{r}
# creating the wine type variable
white_wine <- white_wine |>
  mutate(
    winetype = "white"
  )
```

```{r}

# seeing how it looks
print(white_wine)
```

```{r}
# importing readr, reading in the red wine data

library(readr)
```

```{r}
# making sure to address the semicolon delimiter
red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv", delim = ";")
```

```{r}
# all the column types look good
problems(red_wine)
```

```{r}
# checking what the types are to make sure

spec(red_wine)
```

```{r}
# using the correct column names
colnames(red_wine) <- c(white_wine_variables$Variables)
print(red_wine)
```

```{r}

# creating the wine type variable for the red wine
red_wine <- red_wine |>
  mutate(
    winetype = "red"
  )
```

```{r}
# combine the data sets
new_data <- bind_rows(white_wine, red_wine)
```

```{r}
# checking to see how it looks
print(new_data)
```

```{r}
# Here, I am checking to see that there is in fact red and white data in this new object
new_data |>
  filter(winetype == "red")
```

```{r}
# and also white
new_data |>
  filter(winetype == "white")
```

```{r}
#installing a library
library(dplyr)
```

```{r}
# creating a chain for the new combined data file 
new_data |>
  filter(quality > 6.5 & alcohol < 132) |>
  arrange(desc(quality)) |>
  select(contains("acid"), alcohol, winetype, quality) |>
  mutate(across(!alcohol), mean, .names("mean_alcohol")
)

```
